---
title: "Australian Atlas"
author: "Yashika"
output:
  html_document: default
---
EXAMPLE1
```{r setup}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE
)


```
```{r library, echo=FALSE}
library(readr)
library(tidyverse)
library(galah)

```

## About the data
## Making a report

We have pulled some data on platypuses from the Atlas of Living Australia using the "ALA4R" package, and done some subsetting of the data to focus on just a few of the key variables.

Let's read in data with `read_csv()`:

```{r read-platy, echo=TRUE}
library(readr)
platypus <- read_csv("data/platypus.csv")

```


This code can be read as: "read in this .csv file from the data folder".

There are `r nrow(platypus)` observations of platypuses.


## Creating plots

Let's create some plots. 

Two of the variables in the data set are `latitude` and `longitude` indicating the location where the animal was spotted. This is the first plot, created using the `ggplot2` package from `tidyverse` suite.

```{r plot-platy, echo = FALSE, message=FALSE}
library(tidyverse)
ggplot(data = platypus,
       aes(x = longitude, 
           y = latitude)) + 
  geom_point()
```

## Viewing at the data

You can type `View(platypus)` in the R console to open a "spreadsheet" view of the data.

## Subsetting the data

Let's subset the data to only platypus sightings from Victoria:

```{r filter-platy}
platypus_nsw <- platypus %>% 
  filter(state == "Victoria")
```



## Appendix (This a demo for you to see)

**This is a demo to show you what complex R commands look like. We are going to learn each of the bits and pieces during the semester. This is just a demo so you know what you will be learning soon!**

Below code chunk contains the code used to download the data.

```{r ala-for-4, echo=TRUE, eval = FALSE}
# load the ALA4R package, which contains the platypus data
library(ALA4R)
# Take a look at what the package does using the code `help(package="ALA4R")`

# Look up the scientific name for platypus using:
specieslist("platypus")

# This returns a lot of different organisms with "platypus" in the name, but you should be able to find one line with the relevant information, that its scientific name is "Ornithorhynchus anatinus". 

platypus <- occurrences("Ornithorhynchus anatinus", download_reason_id=10)

# 518426.7 is NOWHERE near Australia. Let's filter it out

# subset the data using dplyr filter and select commands
platypus <- platypus$data %>% 
  filter(longitude < 518426) %>% 
  select(id,
         commonName,
         scientificName,
         state,
         latitude,
         longitude,
         eventDate,
         sex) %>% 
  as_tibble()

write_csv(platypus_nsw, "data/platypus.csv")
```

## Writing CSV file

```{r write-platy-nsw}
# You can learn more about `read_csv` and `write_csv` by typing `?write_csv`.
write_csv(platypus_nsw, "data/platypus_nsw.csv")
```
